---
title: "A2"
author: "Group 10"
date: "9/12/2021"
output: 
  bookdown::pdf_document2:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(visdat)
library(simputation)
library(fastDummies) # does not work for multi-level factors?
```


```{r read_data}
bankruptcy <- readRDS(here::here("data/Bankruptcy.rds"))
```

# Data description

Data are collected on `r ncol(bankruptcy)` variables each representing different measures of status of `r nrow(bankruptcy)` bankrupt companies in the US. These are:

* Name of the firm (`Name`)
* Total assets (in millions of dollars) (`Assets`)
* City where filing took place (`CityFiled`)
* U.S CIP at the time of filing (`CPI`)
* Length of bankruptcy process (`DaysIn`)
* CityFiled, categorized as Wilmington (DE), New York (NY) or all other cities (OT) (`DENYOther`)
* Earnings (operating income) at time of filing (in millions of dollars) (`Ebit`)
* Number of employees before bankruptcy (`Employees`)
* Number of union employees before bankruptcy (`EmplUnion`)
* Total number of other bankrupcy filings in the year of this filing (`FilingRate`)
* Short description of the event that ended the firm’s existence (`FirmEnd`)
* Gross Domestic Product for the Quarter in which the case was filed (`GDP`)
* The population of the firms headquarters city (`HeadCityPop`)
* The distance in miles from the firms headquarters city to the city in which the case was filed (`HeadCourtCityToDE`)
* The state in which firms headquarters is located (`HeadStAtFiling`)
* Total amount of money owed (in millions of dollars) (`Liab`)
* Categorical variable where numbers from 1 to 12 correspond to months from Jan to Dec (`MonthFiled`)
* Prime rate of interest on the bankruptcy filing date (`PrimeFiling`)
* Sales before bankruptcy (in dollars) (`Sales`)
* Standard industrial clasification code (`SICMajGroup`)
* Year bankruptcy was filed (`YearFiled`)

Among these variables, `Assets`, `Ebit`, `GDP`, `Liab` and `Sales` are the measures of economic status of the companies. `CPI`, `PrimeFiling` and `CityFiled` describe the external environment of the companies. `FirmEnd` tells three different endings of the companies: merged with others, bankruptcy, and continuing the operations.

```{r vismiss, }
vis_miss(bankruptcy)
```

the lowest rate of interest at which money may be borrowed commercially

```{r}
bankruptcy %>% 
  dplyr::select_if(is.numeric) %>% 
  GGally::ggpairs()
```

Outliers:

credibiliy:

```{r}
summary(bankruptcy)
```

# Imputation

```{r}
impute_lm(bankruptcy, Liab ~ Assets) %>% 
  impute_lm(Ebit ~ Assets) %>%
  impute_knn(Sales ~ Assets + Ebit + Liab, pool = "univariate", k = 5) %>% 
  impute_knn(Employees ~ Assets + Ebit + Sales, pool = "univariate", k = 5) %>% 
  # impute_lm(DaysIn ~ Assets) %>% 
  separate(SICMajGroup, c("group_code", "group_name"), " ", extra = "merge") %>% 
  mutate(group_code = as.numeric(group_code))-> bankruptcy_imp

bankruptcy_clean <- bankruptcy_imp %>% 
  mutate(group_code = group_code - min(bankruptcy_imp$group_code),
         group_code = as.factor(group_code))
```

`bankruptcy_clean` is the data set after imputation.

# PCA

```{r}
pcaout <- bankruptcy_clean %>% 
  mutate(Name = abbreviate(Name)) %>% 
  column_to_rownames('Name') %>% 
  dplyr::select_if(is.numeric) %>% 
  select(-DaysIn, -EmplUnion, -HeadCourtCityToDE) %>% 
  # select(Assets, CPI, FilingRate, GDP, HeadCityPop, MonthFiled, PrimeFiling, YearFiled) %>%
  prcomp(scale. =TRUE)
```

```{r}
summary(pcaout)
```

```{r}
screeplot(pcaout, type = 'l')
```


```{r}
biplot(pcaout, scale = 0)
```

outliers:




















